---
title: 'Dataquest Guided Project: Investigating COVID-19 Virus Trends'
author: "Cindy Zhang"
output: 
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    # number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
---
## Introduction

This is my solution to Dataquest's COVID-19 Guided Project from Course 2 (Data Structures in R). It answers the question: **Which countries have had the highest number of positive cases against the number of tests?** 

More details, such as descriptions for variables, can be found in the "ReadMe" file of this project's [repository in GitHub](https://github.com/chundychang/DQ-COVID-project).

```{r setup, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, results = 'markup', message = F)
# knitr::opts_chunk$set(include = F)
# knitr::opts_chunk$set(echo = TRUE)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r pkg, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(tibble)
library(DT)
```

## Loading the Data

```{r dataload}
covid_df <- data.frame(read.csv("covid19.csv"))
dim(covid_df)
vector_cols <- colnames(covid_df)
vector_cols
head(covid_df)
glimpse(covid_df)
```

After downloading the `covid19.csv` file from Dataquest, I loaded the file and determined the dimension of the dataframe. I stored the column names as `vector_cols`, a `r typeof(vector_cols)` data structure. The glimpse() function is useful when exploring a new dataset because it makes it possible to see every column in a data frame and shows as much as it can so that you can visualize what the dataset looks like. 

## Isolating the Necessary Rows

As shown in the glimpse of the dataset above, the `Province_State` column mixes country and state/province level data. I was only interested in country-level data, so I filtered `covid_df` for rows that contain `"All States"` in the `Province_State` column and stored it as `covid_df_all_states`. I then deleted `Province_State` from the new dataframe with the assurance that the new dataframe only contained country-level data.

```{r filter}
covid_df_all_states <- covid_df %>%
  filter(Province_State=="All States")
covid_df_all_states$Province_State <- NULL
```

## Isolating the Necessary Columns

From `covid_df_all_states`, I extracted another subset, `covid_df_all_states_daily`, which contained only the dataset columns with daily information to avoid biasing the analysis by comparing daily data to cumulative data. 

```{r dailymeasure}
covid_df_all_states_daily <- subset(covid_df_all_states, select = c(Date, Country_Region, active, hospitalizedCurr, daily_tested, daily_positive))
```

## Extracting Top Ten Tested Cases Countries

I calculated the sum of COVID-109 tested, positive, active, and hospitalized cases by country using the daily data; from the sum calculations, I extracted the top ten countries with the highest sums.

```{r dailysum}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>%
  group_by(Country_Region) %>%
  summarize(tested = sum(daily_tested), positive = sum(daily_positive), active = sum(active), hospitalized = sum(hospitalizedCurr)) %>%
  arrange(desc(tested))
covid_df_all_states_daily_sum
covid_top_10 <- head(covid_df_all_states_daily_sum, 10)
covid_top_10
```

## Identifying the Highest Positive Against Tested Cases

I extracted vectors from `covid_top_10` that allowed me to calculate the ratio of positive cases to tested cases and determine which three countries had the highest ratios.

```{r vector}
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized
```

```{r vectornames}
names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries
```

```{r topthree}
positive_tested_ratio <- sort(positive_cases/tested_cases, decreasing=TRUE)
positive_tested_top_3 <- positive_tested_ratio[1:3]
```

## Keeping Relevant Information

I created a matrix that contained just the top three countries' COVID-19 information:

```{r vectortomatrix}
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)
covid_mat <- rbind(united_kingdom, united_states, turkey)
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")
covid_mat
```

## Putting It All Together

Lastly, I stored all answers and datasets together in one list, `covid_analysis_list`. The second element of this list is printed below the code:

```{r list}
question <- "Which countries have had the highest number of positive cases against the number of tests?"
answer <- c("Positive tested cases" = positive_tested_top_3)
dataframes <- c(covid_df, covid_df_all_states, covid_df_all_states_daily, covid_df_all_states_daily_sum, covid_top_10)
matrices <- covid_mat
vectors <- c(active_cases, countries, hospitalized_cases, positive_cases, positive_tested_ratio, positive_tested_top_3, tested_cases, turkey, united_kingdom, united_states, vector_cols)
data_structure_list <- c(dataframes, matrices, vectors)
covid_analysis_list <- c(question, answer, data_structure_list)
covid_analysis_list[2]
```